---
title: "Prediction Assignment1"
author: "Ning Guo"
date: "2016/12/2"
output: html_document
---

#Synopsis#
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Loading the data#
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
```{r}
library(caret); library(ggplot2); library(randomForest)
fileURL1<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fileURL2<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if(!file.exists("training")|!file.exists("testing")){
        # create the placeholder file
        td = tempdir()
        # download into the placeholder file 
        tf1 = tempfile(tmpdir=td)
        download.file(fileURL1, tf1)
        pmltraining<- read.csv(tf1)
       
        tf2 = tempfile(tmpdir=td)
        download.file(fileURL2, tf2)
        pmltesting<- read.csv(tf2)     
}
dim(pmltraining)
dim(pmltesting)
```


#Data pre-processing# 
```{r}
## Removing near zero variables
nzv<- nearZeroVar(pmltraining)
pmltrainingtemp<- pmltraining[,-nzv]
## Removing columns mosting with NAs
threshod<- dim(pmltraining)[1]*0.9
badcolumes<- which(apply(is.na(pmltrainingtemp), 2, sum) > threshod)
pmltrainingtidy<- pmltrainingtemp[,-badcolumes]
pmltestingtemp<- pmltesting[,-nzv]
pmltestingtidy<- pmltestingtemp[, -badcolumes]
## Removing non-measurement data
RemoveInx1 <- grepl("X|timestamp|user_name|problem_id", names(pmltrainingtidy))
RemoveInx2 <- grepl("X|timestamp|user_name|problem_id", names(pmltestingtidy))
pmltrainingtidy<- pmltrainingtidy[, which(RemoveInx1==FALSE)]
pmltestingtidy<- pmltestingtidy[, which(RemoveInx2==FALSE)]
```

#Data splitting for resampling# 
```{r}
set.seed(123)
inTrain<- createDataPartition(y=pmltrainingtidy$classe, p=0.7, list = FALSE)
training<- pmltrainingtidy[inTrain,]
testing<- pmltrainingtidy[-inTrain,]
```

#Model fitting#
I tried three different methods for modeling, including random forest, decision tree, LDA
```{r}
## random forest
modfit1<- randomForest(classe~., data = training)
prediction1<- predict(modfit1, testing)
confusionMatrix(testing$classe, prediction1)
## decision tree
## modfit2<- train(classe~., method="rpart", data = training)
## prediction2<- predict(modfit2, testing)
## confusionMatrix(testing$classe, prediction2)
## LDA
## modfit3<- train(classe~.,  method="lda", data = training)
## prediction3<- predict(modfit3, testing)
## confusionMatrix(testing$classe,  prediction3)
```
Random forest had the best performance with an accuracy of 99.68%. So random forest was used as the final model.

```{r}
## important variables
impVar<- varImp(modfit1)
impVar$names<- row.names(impVar)
impVar<- impVar[order(impVar[,1], decreasing = FALSE),]
dotchart(impVar$Overall[34:53], labels = impVar$names[34:53], xlab = "varImp", main = "Top20 important variables in random forest modeling")
```

From the docchart, we found that "num_window" "roll_belt" "yaw_belt" were the TOP3 important variables for modelling.

#Predict the test data#
```{r}
predanswer<- predict(modfit1, pmltestingtidy)
predanswer
```

